---
title       : "IRIS Flower Data Set"
subtitle    : "Developing Data Product- Course Project"
author      : "JMandrake"
job         : "Data Science Enthusiast"
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## About IRIS Flower Data Set

- A multivariate data set introduced by Sir Ronald Fisher (1936) as an example of discriminant analysis.
- The data set consists of 50 samples from each of three species of Iris:
        
        1. Iris setosa
        2. Iris virginica
        3. Iris versicolor
- Four features were measured (in cm) from each sample:

        - Sepal length and width
        - Petal length and width
- In this slidify deck we present exploratory data analysis with this data set.

---

## Loading IRIS Data and Summarizing.

```{r}
library(datasets) #Load dataset library
data(iris) #Load iris dataset
summary(iris)
```

---

## Box Plot for Different Species

```{r, echo=FALSE}
par(mfrow=c(2, 2))
boxplot(Sepal.Length~ Species, data = iris, main = "Sepal Length vs Species" )
boxplot(Sepal.Width~ Species, data = iris, main = "Sepal Width vs Species")
boxplot(Petal.Length~ Species, data = iris, main = "Petal Length vs Species")
boxplot(Petal.Width~ Species, data = iris, main = "Petal Width vs Species")
```

---

## Scatter Plot for different species

```{r, echo=FALSE}
pairs(iris[,1:4],col=iris$Species, main = "Iris Flower Data Set")
```




